---
title: "Sampling demonstration"
author: "Lukáš Hejtmánek"
date: "8 October 2018"
output: github_document
---

```{r setup, include=FALSE}
library(ggplot2)
library(dplyr)
library(knitr)
set.seed(100)
N <- 1000
normal <-  rnorm(N, 100, 15)
normal_shifted <- rnorm(N, 110, 15)
extremes <- c((15 * rnorm(N/2, 0, 1)) + 50, (15 * rnorm(N/2, 0, 1)) + 150)
bias_extremes <- c(rnorm(N/4, 75, 5), c(rnorm(N*3/4, 110, 5)))
missing <- c(normal[sample(1:N,round(0.7*N), replace=F)],rep(NA, round(0.3*N)))

types <- c(rep("normal", N), rep("normal_shifted", N), 
           rep("extremes", N), rep("bias_extremes", N), 
           rep("missing", N))

values <- c(normal, normal_shifted, extremes, bias_extremes, missing)
mean(values, na.rm = T)

df <- data.frame(value=values, type=types)
  
sample_data <- function(data, n_sample){
  selected <- c()
  for(i in 0:4){
    selected <- c(selected, sample(1:N, n_sample) + N*i)
  }
  sampled_data <- data[selected,]
  ls <- list()
  ls$mean <- mean(sampled_data$value, na.rm = T)
  ls$mean_plot <- ggplot(sampled_data, aes(type, value, fill=type)) + stat_summary(fun.y="mean", geom="bar") + 
    stat_summary(geom = "errorbar", fun.data = mean_se, position = "dodge", width=0.2) + 
    guides(fill=FALSE)
  ls$hist_plot <- ggplot(sampled_data, aes(value, fill=type)) + geom_histogram(bins=50) + geom_density(aes(y=..density..*5)) + 
    facet_wrap(~type) + geom_vline(xintercept = ls$mean) + guides(fill=FALSE)
  ls$data <- sampled_data
  return(ls)
}
ls_all <- sample_data(df, N)
```

# Basic descriptors

So in our case we have a mean `r ls_all$mean` and a following distribution.

```{r, echo=F, warning=F}
ls_all$mean_plot
true_mean <- ls_all$data %>% group_by(type) %>%  summarize(true_mean=mean(value, na.rm = T))
```

looking at each separately, we get the following distribution. ALL have approximate mean of 100, except normal shifted value.

```{r, echo=F, message=F, warning=F}
ls_all$hist_plot
```

## Sampling
### Sampling 10 percent of the sample
Sampling ten percent values. We will see what happens if we do that 100 times
```{r, echo=F, message=F, warning=F}
ls_10 <- list()
n_sample <-  round(N/10)
df_10_all_data <- data.frame(value=c(), type=c(), sample=c())
for(i in 1:100){
  ls_10[[i]] <- sample_data(df,n_sample)
  new_data <- ls_10[[i]]$data
  new_data$sample <- rep(i, n_sample)
  df_10_all_data <- rbind(df_10_all_data, new_data)
}
```

Let's have a look at the means of those 100 samples of 10 percent of data
```{r, echo=F, message=F, warning=F}
means_10 <- df_10_all_data %>% group_by(sample, type) %>% summarize(mean=mean(value, na.rm=T))
ggplot(means_10, aes(mean, fill=type)) + geom_histogram(bins=30) + facet_wrap(~type) + guides(fill=F)
range_10 <- means_10 %>% group_by(type) %>% summarize(min=min(mean, na.rm = T), max = max(mean, na.rm=T), mean=mean(mean, na.rm=T))
range_10 <- merge(range_10, true_mean, by="type")
kable(range_10)
```


## Sampling 5 percent

```{r, echo=F, message=F, warning=F}
ls_5 <- list()
n_sample <-  round(N/20)
df_5_all_data <- data.frame(value=c(), type=c(), sample=c())
for(i in 1:100){
  ls_5[[i]] <- sample_data(df,n_sample)
  new_data <- ls_5[[i]]$data
  new_data$sample <- rep(i, n_sample)
  df_5_all_data <- rbind(df_5_all_data, new_data)
}
```

Let's have a look at the means of those 100 samples of 10 percent of data
```{r, echo=F, message=F, warning=F}
means_5 <- df_5_all_data %>% group_by(sample, type) %>% summarize(mean=mean(value, na.rm=T))
ggplot(means_5, aes(mean, fill=type)) + geom_histogram(bins=30) + facet_wrap(~type) + guides(fill=F)
range_5 <- means_5 %>% group_by(type) %>% summarize(min=min(mean, na.rm = T), max = max(mean, na.rm=T), mean=mean(mean, na.rm=T))
range_5 <- merge(range_5, true_mean, by="type")
kable(range_5)
```
